---
title: "Course Project - Regression Models"
author: "Venkat Ram Rao"
date: "11/22/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This report represents the final course project for the Regression Models course, part of the John Hopkins  Statistics and Machine Learning Specialization on Coursera

The purpose of the report is to explore the relationship between a set of variables and miles per gallon (MPG) (outcome). In particular answer the following questions:

1. “Is an automatic or manual transmission better for MPG”
2. "Quantify the MPG difference between automatic and manual transmissions"

## Exploratory analysis
mtcars is a dataset provided by R and contains 32 sample records from 1974 comparing fuel consumption(mpg) and 10 aspects of automobile design and performance for 32 automobiles 
```{r mtcars}

data(mtcars)
dim(mtcars)
str(mtcars)
```
Below is a simple plot showing the miles/gallon for manual and automatic transmission.
```{r}
library(ggplot2)
my_boxplot <- ggplot(mtcars,aes(x=am,y=mpg, group=am, col=am)) + geom_boxplot() + xlab('Transmission (0 = automatic, 1 = manual)') + ylab('Miles per Gallon')
my_boxplot 
```
At first glace, there seems to be a clear difference in mpg based on the Transmission.

A better analysis can be done by running a T-test
```{r}
mtcars_am0 <- mtcars[which(mtcars$am == 0), names(mtcars) %in% c("mpg")]
mtcars_am1 <- mtcars[which(mtcars$am == 1), names(mtcars) %in% c("mpg")]
t.test(mtcars_am0, mtcars_am1, paired = FALSE, alternative="two.sided", var.equal=FALSE)
```
This shows that there is a mean difference of 7.24492mpg between cars with Manual versus Automatic transmission. The 95% confidence interval (-11.28,-3.21)  does not contain zero meaning we can say with a 95% confidence that Manual Transmission leads to better fuel economy.

However, there are other variables which effect the mpg so additional analysis will need to be done as below

## Regression Models

#Data Prep:

It seems all the data in mtcars is defined as numeric. I re-factored the 'am' column to better reflect the data:
```{r}
mtcars$am<-relevel(as.factor(mtcars$am), ref='0',labels=c('Automatic','Manual'))
```

#Simple model:

Starting with a simple model of mpg vs Transmission type

```{r}
lm_am <- lm(mpg~factor(am) -1,data = mtcars)
summary(lm_am)
```
shows a model with a high R-squared of (.9487). However, plotting the residuals:
```{r}
plot(predict(lm_am),resid(lm_am))
```
shows that the residual plot is definitely not random. Hence, a simple regression model with just Transmission type of not reliable.

#Best model:
Running a regression with all the variables show multiple variables with a strong correlation with mpg
```{r}
lm_all <- lm(mpg~. -1,data = mtcars)
summary(lm_all)
```
In particular, weight(wt) has a strong negative coefficient. 1/4 mile time (qsec), Number of forward gears(gear) and Rear axle ratio(drat) seem to have a strong positive correlation as well

The best model can be found by using the step function
```{r}
lm_best<-step(lm_all,direction ="both")
summary(lm_best)
```
This gives us a model based on wt, qsec and tramsmission.
wt    -3.9165     0.7112  -5.507 6.95e-06 ***
qsec   1.2259     0.2887   4.247 0.000216 ***
am0    9.6178     6.9596   1.382 0.177915    
am1   12.5536     6.0573   2.072 0.047543 *  

```{r}
par(mfrow = c(2,2))
plot(lm_best)
```
The residual plot is fairly random as well.

##Conclusion

Looking at the coefficients, it looks like:
1. Changing the transmission from Automatic to Manual causes an increase in Fuel economy of 2.9358 mpg when controlled for all factors. This is less than what we got from a simple T-test but more accurate given the multitude of other factors.
2. Weight and 1/4 mile time are also influential. Each 1000 lb increase in weight reduces fuel economy by 3.9165 mpg, 
3. Increasing qsec also increases  Fuel Economy. This is interesting as there is no logical reason for this to be the case. This needs to be investigated further. There is a possibility
that some bias crept in and qsec has strong correlations with other factors (e.g. hp)

##Assumptions:
1. The big issue with this analysis is that there are only 32 records in the data set. This is fairly small and if posible additional data needs to be identified.